Bootstrap learning via modular concept discovery

### 2013-04-23 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Added Josh's version of the introduction.

\section{Introduction}
A central question for AI is how to create intelligent agents without
building in vast amounts of pre-specified knowledge that is tailored
to one domain or another. This is generally accomplished by a learning
algorithm that uses past experience to improve future performance. The
problem is that if past problem solving attempts entirely missed the
mark, their failures are unlikely to shed light on will work and what
will not. Such uninformative solutions, however, vastly outnumber the
rest, and our uninformed agent has no way to know which solutions will
be informative and which will not. Thus, our learning algorithms must
contend with a bootstrapping problem: the less an agent knows to begin
with, the harder it is to learn from experience.

For example, imagine you know nothing about electronics but are
supplied with a collection of circuit elements and asked to build a
number of canonical useful devices, such as a digital adder, an
op-amp,, etc. Your initial attempts will be uniformly ineffective. If,
on the other hand, you know the basics of circuit design, you are far
more likely to generate devices that are somewhat successful and on
whose basis you might improve your designs and your design skills.

Like many of the problems in AI, this one can be cast as a search
problem: the goal is to find a circuit wiring procedure that maximizes
certain criteria.  There are many search algorithms, but they typically
exploit local or topological structure in the search space. For
example, local search algorithms move from one candidate solution to a
neighboring one. Heuristic search techniques take advantage of
topological structure to prune off large parts of the solution space.

This structure derives from domain specific knowledge. In local
search, domain knowledge provides the similarity metric between
solutions; in heuristic search it provides the heuristic. Here we ask
how to bootstrap domain specific knowledge and extract structure from
the space of possible solutions by attempting to solve multiple
related problems in a given domain at once. Our approach is motivated
by the observation that many interesting problems -- such as
discovering scientific theories, designing new technologies, learning
a new language or learning a skill such as playing an instrument or
cooking -- have solutions that are naturally specified as programs.

Our proposed solution is thus inspired by the programmer who -- having
reused subroutines to solve similar simple tasks -- creates a function
library to help manage the complexity of larger projects. The
E.C. algorithm, which we introduce here, builds a library of reusable
program components and places a distribution over these, effectively
learning from simple tasks a search space structure that enables it to
solve more complex tasks.

>>>>

\section{Introduction}

Imagine that you know nothing about electronics but are supplied with
a collection of circuit elements and asked to build a number of useful
digital devices, such as an adder, a flip-flop, a counter, a
multiplexer, and so on.  Your initial attempts will probably be
uniformly ineffective.  If your only feedback is based on success or
failure, you may have trouble learning anything.  If, on the other
hand, you already know the basics of circuit design, you are far more
likely to generate some devices that are at least somewhat successful.
In this case, with feedback you can improve your designs and your
design skills.

This example illustrates the central ``bootstrapping'' challenge for
attempts to create intelligent agents without building in vast amounts
of pre-speciﬁed domain knowledge.  Learning from experience is
typically most effective in a system that already knows most of what
it needs to know, because mistakes are most informative.  In a system
that knows almost nothing, where most problem solving attempts
entirely miss their mark, failures are unlikely to shed light on will
work and what will not.  Uninformative solutions vastly outnumber the
rest, and an uninformed agent has no way to know which solutions will
be informative and which will not.  If it is harder to learn the less
an agent knows, how can a learner who knows almost nothing about a domain
ever get off the ground?

Here we propose an answer to this challenge, based on an approach to
discovering new and increasingly sophisticated concepts that build
modularly on previously learned, simpler concepts.  Like many problems
in AI, the bootstrapping challenge and its solution can be cast in
terms of search.  In our circuit-building scenario, the goal is to ﬁnd
circuit wiring procedures that maximize certain criteria.  Search
algorithm typically aim to exploit local or topological structure in
the search space -- here the space of all circuit-wiring procedures.
For example, local search algorithms move from one candidate solution
to a neighboring one. Heuristic search techniques take advantage of
topological structure to prune off large parts of the solution space.
This structure derives from domain-speciﬁc knowledge.  In local
search, domain knowledge provides the similarity metric between
solutions; in heuristic search it provides the heuristic.  Here we ask
how to bootstrap domain-speciﬁc knowledge by extracting search
structure from the space of possible solutions, by attempting to solve
multiple related problems in a given domain at once.

Our approach is motivated by the observation that many interesting
problems –- such as discovering scientiﬁc theories, designing new
technologies, learning a skill such as playing an instrument or
cooking –- have solutions that are naturally speciﬁed as programs.
Our proposed solution is inspired by the programmer who, having found
various subroutines that can be reused to solve similar simple tasks,
creates a function library to help manage the complexity of larger
projects. The E.C. algorithm, which we introduce here, builds a
library of reusable program components and places a distribution over
these, effectively learning from simple tasks a search space structure
that enables it to solve more complex tasks.
